---
title: "Developing Data Products"
subtitle: "U.S. SARS-COV-2 Map Application"
author: "Eric Smith"
date: "7/21/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

source("scripts/strsplit_i.R")
source("scripts/get_report.R")
source("scripts/prep_combined_key.R")
source("scripts/calc_Active.R")
source("data/geo_data/state_codes.R")
library(dplyr)
library(ggplot2)
```

```{r covid data, message=FALSE}
report_dir <- "covid_data/county_reports/"
file_names <- list.files(report_dir,
                         full.names = TRUE)

## Compile COVID reports into one data frame
reports <- do.call(rbind,
                   lapply(file_names, get_report))

## Summarize data into US and State totals by date
us_totals <- reports %>%
        select(Date, Deaths, Confirmed, Recovered) %>%
        group_by(Date) %>%
        summarize(Deaths = sum(Deaths),
                  Confirmed = sum(Confirmed),
                  Recovered = sum(Recovered))
st_totals <- reports %>%
        select(Date, State, Deaths, Confirmed, Recovered) %>%
        group_by(Date, State) %>%
        summarize(Deaths = sum(Deaths),
                  Confirmed = sum(Confirmed),
                  Recovered = sum(Recovered))

st_totals$Active <- with(st_totals, 
                         calc_Active(Confirmed, Deaths, Recovered))
reports$Active <- with(reports, 
                       calc_Active(Confirmed, Deaths, Recovered))

```

```{r pop data, message=FALSE}
data_fld <- "data/population_data/"

## Read population data and select relevant variables:
pop_data <- read.csv(paste0(data_fld, "co-est2019-alldata.csv"))
pop_data <- pop_data %>%
        transmute(CTYNUM = COUNTY,
                  State = STNAME,
                  County = CTYNAME,
                  Population = POPESTIMATE2019)

## Divide population data set into county and state data sets
county_pop <- pop_data[pop_data$CTYNUM != 0, c("County", 
                                               "State", 
                                               "Population")]
state_pop <- pop_data[pop_data$CTYNUM == 0, c("State", 
                                              "Population")]

## Create a common key to combine pop data with county level covid data
county_pop$Combine_Key <- apply(county_pop[,c("County","State")], 
                                1, 
                                prep_com_key)

reports$Combine_Key <- apply(reports[,c("County","State")],
                             1,
                             prep_com_key)

## Merge covid and population data sets
county_data <- merge(reports,
                     county_pop[,c("Combine_Key",
                                   "Population")])
state_data <- merge(st_totals,
                    state_pop)

```

```{r final prep}
## Divide by population to find per capita indicators
state_data$Deaths.per.capita <- state_data$Deaths / state_data$Population
state_data$Confirmed.per.capita <- state_data$Confirmed / state_data$Population
state_data$Active.per.capita <- state_data$Active / state_data$Population

county_data$Deaths.per.capita <- county_data$Deaths / county_data$Population
county_data$Confirmed.per.capita <- county_data$Confirmed / county_data$Population
county_data$Active.per.capita <- county_data$Active / county_data$Population

## Add Name variable to display on hover
state_data$Name <- state_data$State
county_data$Name <- county_data$County

## Add state codes for selecting states to map
State <- names(State_Codes)
county_data <- merge(county_data, cbind(State,State_Codes))

## Filter county data to relevant columns for map application
county_data <- county_data[,c("Name","State_Codes","LatLong","Population","Date",
                              "Confirmed","Deaths","Active",
                              "Confirmed.per.capita","Deaths.per.capita","Active.per.capita")]

```

```{r write data sets, eval=FALSE}
write.csv(county_data, "data_sets/county_data.csv")
write.csv(state_data, "data_sets/state_data.csv")
```
## BACKGROUND
### SARS-COV-2 in the United States

SARS-COV-2 is the name for the coronavirus that causes the deadly disease COVID-19.
Its spread through the United States began in early 2020 and continues today.
The prevalence and spread of the virus in the country have varied widely from
region to region due, in part, to political and demographic differences between States.
Comparing trends by region will be an important strategy in making effective
policy decisions throughout the remainder of the outbreak.

## BACKGROUND


```{r, fig.height=3.25}
g_us <- ggplot(us_totals, aes(x = Date, y = Deaths / 1000))
g_us + geom_line(size = 1.5, color = hcl(h=195,l=65,c=100)) +
        ylab("Thousands of Deaths") + xlab("") + 
        ggtitle("Cumulative Coronavirus Deaths in the U.S.") +
        theme(text = element_text(size = 17), 
              axis.text.x = element_text(angle = 30)) +
        scale_x_date(date_breaks = "3 weeks",
                     date_labels = "%b-%d")
```

Despite efforts to curb the impact of the virus, deaths continued to climb steadily throughout June and July. The toll has now nearly reached 150,000 deaths.

## PROJECT
### U.S. COVID-19 Map Application

The goal of this project is to create an R Shiny App with
the following features:

- A "snapshot" of the virus' spread at a date selected by the user
- Ability to toggle between *total* and *per capita* cases or deaths
- Adjustable scale such that the user can view the entire U.S. or sub-regions

The app is available [here](https://ericeliassmith.shinyapps.io/Course_Project/).

## DATA

The coronavirus indicator data for the project came from the
[COVID-19 Data Repository](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_daily_reports) 
maintained by the Johns Hopkins University Center for Systems Science and
Engineering (JHU CSSE).
JHU CSSE uses this data to operate the 
[2019 Novel Coronavirus Visual Dashboard](https://www.arcgis.com/apps/opsdashboard/index.html#/bda7594740fd40299423467b48e9ecf6).

The population data for the project came from the
[United States Census Bureau](census.gov).
Specifically, 2019 population estimates were used from the data set available
[here](https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv).

The code used to process these data sets is embedded in the R Markdown file that produced this presentation. It is available [here](https://github.com/1earning-R/Dev_Data_Products_Project_COVID/blob/master/pitch.Rmd).